The Weakly Nonlinear Magnetorotational Instability

### 2016-08-15 paper/widegap_mri_paper.tex
widegap added description of adjoint to appendix

\begin{widetext}
\begin{table*}[ht]
\normalsize
\caption{Fiducial parameters for MRI runs} \label{table:parameters}
\centering
%\resizebox{\textwidth}{!}
{\begin{tabular}{rlrrrrrrr}
  \hline
 & $\xi$ & $Pm$ & $\beta$ & $\Omega_2/\Omega_1$ & $R_1/R_2$ & radial magnetic b.c. \\ 
  \hline\hline
Standard MRI & 0 & 1E-3 & 25 & 0.18 & 0.33 & conducting \\ 
Helical MRI & 4 & 1E-6 & 1.7E-2 & 0.27 & 0.5 & insulating\\ 
   \hline
\end{tabular}}
\end{table*}

@@

>>>>

\begin{widetext}
\begin{table*}[ht]
\normalsize
\caption{Fiducial parameters for MRI runs} \label{table:parameters}
\centering
%\resizebox{\textwidth}{!}
{\begin{tabular}{rlrrrrrrr}
  \hline
 & $\xi$ & $Pm$ & $\beta$ & $\Omega_2/\Omega_1$ & $R_1/R_2$ & radial magnetic b.c. \\ 
  \hline\hline
%Standard MRI & 0 & 1E-3 & 25 & 0.18 & 0.33 & conducting \\ 
Standard MRI & 0 & 1.6E-6 & 41.2 & 0.121 & 0.33 & conducting \\ 
Helical MRI & 4 & 1E-6 & 1.7E-2 & 0.27 & 0.5 & insulating\\ 
   \hline
\end{tabular}}
\end{table*}

@@

%The GLE is derived from 
%Equation \ref{eq:ordere3} is resonant with 
To find a bounded solution at $\mathcal{O}(\epsilon^3)$ we must eliminate secular terms: terms which are resonant with the solution to the homogenous equation $(\mathcal{L} + \xi \widetilde{\mathcal{H}}) \mathbf{V} = 0$ and cause the solution to grow without bound. To eliminate these terms we enforce a solvability condition, which arises as a consequence of the Fredholm alternative. This solvability condition is the requirement that the inner product of the equation with its adjoint homogenous solution is zero. We define the adjoint operator $\mathcal{L}^\dagger$ and solution $\mathbf{V}^\dagger$ as

\beq
\langle \mathbf{V^\dagger} | (\mathcal{L} + \xi \widetilde{\mathcal{H}}) \mathbf{V} \rangle = \langle (\mathcal{L}^\dagger + \mathcal{H}^\dagger) \mathbf{V}^\dagger | \mathbf{V} \rangle,
\eeq

where the inner product is defined as

\beq\label{eq:inner_product_def}
\langle \mathbf{V^\dagger} | \mathcal{L} \mathbf{V} \rangle = \frac{k_c}{2\pi} \int_{-\pi/k_c}^{\pi/k_c} \int_{r_1}^{r_2} \mathbf{V}^{\dagger*} \cdot \mathcal{L} \mathbf{V} \, \mathrm{d} r \mathrm{d} z
\eeq

We find that

\beq
\mathcal{L}^\dagger = \mathcal{L}_0^\dagger - \partial_z \mathcal{L}_1^\mathrm{T} + d_z^2 \mathcal{L}_2^\mathrm{T} - \partial_z^3 \mathcal{L}_3^\mathrm{T} + \partial_z^4 L_4^\mathrm{T},
\eeq

\beq
\mathcal{H}^\dagger = - d_z \mathcal{H}^\mathrm{T}
\eeq

where

\beq
\mathcal{L}_0^\dagger = \left[\begin{matrix}
-\frac{1}{\reye} (-\frac{3}{r^4} \partial_r + \frac{3}{r^3}\partial_r^2 - \frac{2}{r^2}\partial_r^3 + \frac{1}{r}\partial_r^4) & 0 & 0 & 0 \\
0 & -\frac{1}{\reym} (\partial_r^2 - \frac{1}{r} \partial_r) & 0 & 0 \\
0 & 0 &-\frac{1}{\reye} (\partial_r^2 + \frac{1}{r}\partial_r - \frac{1}{r^2}) & 0 \\
0 & 0 & 0 & -\frac{1}{\reym} (\partial_r^2 - \frac{1}{r} \partial_r) \\
\end{matrix}\right].
\eeq

The adjoint boundary conditions are selected to satisfy Equation \ref{eq:inner_product_def}, and differ depending on the boundary conditions enforced on the original system. For the conducting boundary conditions we apply to the standard MRI, the adjoint equation

\beq
(\mathcal{L}^\dagger + \xi \widetilde{\mathcal{H}}^\dagger)\mathbf{V}^\dagger = 0
\eeq

must be solved subject to the boundary conditions

\beq
\Psi^\dagger = \partial_r \Psi^\dagger = u^\dagger = A^\dagger = \partial_r B^\dagger = 0.
\eeq
